Associative Data Structures for Website Filtering

### 2014-04-17 paper/paper.tex

    \section{Plan}
    \label{sec:plan}

@@

    \subsection*{Data}
    For our data, we took the system's dictionary file and the list of top level domains from Wikipedia
    and combined these to generate a list of items in the form \texttt{www.\$word.\$tld}, where
    \texttt{\$word} is one of words from the the dictionary and \texttt{\$tld} is a top level domain.  Only
    words consisting of the lowercase letters \texttt{a} through \texttt{z} were used.  Constraining the
    words to consist of 8 to 12 characters and limiting the top-level domains to exactly two letters, 
    all possible combinations were generated.  For the full paths, a number of domains were selected
    and each was appended a random alphanumeric string of eighty characters.  The exact implementation
    can be viewed in the sourcecode\footnote{ref to apendix or code dump}.

@@

    Our secondary point of research encompassed the effect of key complexity on associative data structures.
    Our different types of keys were IP adresses, domain names and full path URLs. We have seen that as expected,
    IP adresses as keys provide a significant improvement over the other two, and uses less memory as well. Interesting to
    note is that we found the only advantage of the binary search tree over the hashmap here, is that insertions of
    a small amount of IP adresses, up to around $2^13$ insertions, is faster. So when writing an application with very
    few keys of low complexity, and with more insertions than queries, the binary search tree may be the best option. Finally,
    we have found that the domain key type does not outperform the full path key type. It is not significantly better
    if it is better at all.
    %The skip list uses the least memory but is awful in every other imaginable way.  The hashmap is
    %usually the best, except when we're dealing with a small number of integer keys and doing far more
    %inserts than queries.  Note that this doesn't actually happen in practice.

>>>>

    \section{Research Plan}
    \label{sec:plan}

    \subsubsection*{Scope and assumptions}

    We make several assumptions for our research.

    Firstly we plan on only working with datasets that are certain to fit into the main memory of our machine. We intend to
    purely examine the relative speed and memory-efficiency of the different data structures. Using datasets larger than
    memory will cause regular lookups to the harddrive, which we assume would skew our results.

    In order to make the measurements more predictable, we will always ensure that insertions are only performed with
    non-existent keys. The alternative would make either the memory usage or number of insertions non-deterministic, which
    could cause inaccuracies in measurement. Similarly, we require that all queries be performed on existing keys.

    To mimic real usage, we will insert and then access the keys in an arbitrary order.  However, we will keep the order consistent
    between tests on different data structures.  Together this will minimize any bias related to any kind of order-dependent
    performance, which would likely be an advantage specific to the implementation, not to the data structure. We will
    generate the order for both lookup and insertion by shuffling the dataset we chose for the specific test. This means that
    orders will not be persistent between different data set sizes; however, we do not expect this to be significant.

    We realize that methods to measure the amount of time required for actions will themselves need time. We will however
    use the same time-measuring method for all data structures, so we can still use the results to compare the different
    data structures.

@@

    \subsubsection*{Data}
    For our data, we used datasets containing $2^x$ elements with $x$ an integer such that $8 < x <
    21$. Thus we have twelve different dataset sizes. We took the system's dictionary file and the
    list of top level domains from Wikipedia and combined these to generate a list of items in the
    form \texttt{www.\$word.\$tld}, where \texttt{\$word} is one of words from the the dictionary
    and \texttt{\$tld} is a top level domain.  Only words consisting of the lowercase letters
    \texttt{a} through \texttt{z} were used. Constraining the words to consist of 8 to 12 characters
    and limiting the top-level domains to exactly two letters, all possible combinations were
    generated.  For the full paths, a number of domains were selected and each was appended a random
    alphanumeric string of eighty characters.  Finally, for the third type of key, an IP address, we
    used 32 bit integers. The exact implementation can be viewed in the sourcecode\footnote{ref to
    apendix or code dump}.

@@

    Our secondary point of research encompassed the effect of key complexity on associative data
    structures.  Our different types of keys were IP adresses, domain names and full path URLs. We
    have seen that as expected, IP adresses as keys provide a significant improvement over the other
    two, and uses less memory as well. Interesting to note is that we found the only advantage of
    the binary search tree over the hashmap here, is that insertions of a small amount of IP
    adresses, up to around $2^13$ insertions, is faster. So when writing an application with very
    few keys of low complexity, and with more insertions than queries, the binary search tree may be
    the best option. Finally, we have found that the domain key type does not outperform the full
    path key type. It is not significantly better if it is better at all.
